# dplyr data manipulation


```{r libraries, message=FALSE, warning=FALSE}
library(PKPDmisc)
library(knitr)
library(tidyverse)
```

Objectives:

* Import datasets and documents
* Perform basic data manipulation upon importing the data.

### Task-I

Use the .csv files `demog`, `IV`, and `Oral` provided into the data object folder. 

1. Read in all three csv files  and give them descriptive names (not data1, data2, data3)

```{r}
demog <- read_csv("../data/demog.csv")
iv_data <- read_csv("../data/IV.csv")
oral_data <- read_csv("../data/ORAL.csv")
```

## DATA MANIPULATION
The goals of this section:

* Use data manipulation tools to prepare the dataset for analysis

### Task-II
1.  Rename "DV" column as "COBS"

```{r}
iv_data <- iv_data %>% rename(COBS = DV)
oral_data <- oral_data %>% rename(COBS = DV)
```

2. Add a Formulation column and label IV/Oral for each dataset

```{r}
iv_data <- iv_data %>% mutate(FORM = "IV")
oral_data <- oral_data %>% mutate(FORM = "ORAL")
```

3. Appropriately merge the demographics dataset into the IV and Oral dataset
4. Create one integrated dataset with both IV and Oral data.

```{r}
combined_data <- bind_rows(iv_data, oral_data)

## check to see if any ids not in the other
anti_join(combined_data, demog)
anti_join(demog, combined_data)
```

Two individuals do not have any concentration-time data

```{r}
all_data <- left_join(combined_data, demog)
```


5. Perform the following tasks:
    a. Ensure that the following columns are numeric and not text: TIME, COBS, WT, AGE, AMT and DOSEs

```{r}
all_data %>% select(TIME, COBS, WT, AGE, AMT, DOSE) %>% str
```

COBS is a character column, therefore want to find out what character values exist

```{r}
# check what character values are present
unique_non_numerics(all_data$COBS)
```

    b. Change the following:
    c. Create a new column called BQLFLAG which takes a value of "0" if there is a numerical value in CObs and "1" if there is "BQL" in CObs.
    
```{r}
# if don't manually specify to handle NA COBS, will also get NA values for BQLFLAG
all_data <- all_data %>% mutate(BQLFLAG = ifelse(is.na(COBS), 0, 
                                                 ifelse(COBS == "BQL", 1, 0)),
                                COBS = as_numeric(COBS))
```

```{r}
all_data %>% head %>% kable
all_data %>% filter(BQLFLAG ==1) %>% kable
```

    d. Filter the dataset such that you remove all rows where BQLFLAG=1
        i. WT from lb to kg 
        iv. CObs from μg/mL to μg/L

```{r}
f_all_data <- all_data %>% filter(BQLFLAG != 1)
f_all_data_adjunits <- f_all_data %>% mutate(WT = WT/2.2,
                                             COBS = COBS*1000)
```

```{r}
f_all_data_adjunits %>% head %>% kable
```

    e. Create a new column called "GENDER" where:
        i. Female = 0
        ii. Male = 1 
    f. Create a new column called RACEN where:
        i. Caucasian = 0
        ii. Asian = 1
        iii. Black = 2
        iv. Hispanic = 3
    g. Create a new column called "LOGCOBS" where CObs is in the log scale
    h. Create a new column called "USUBJID" - unique subject ID as combination of formulation and ID (hint check out `?interaction`)
    
    i. Remove the following columns
    i. SEX
    ii. RACE

```{r}
final_data <- f_all_data_adjunits %>% mutate(
  GENDER = ifelse(SEX == "Female", 0, 1),
  RACEN = as.numeric(factor(RACE, levels = c("Caucasian", "Asian", "Black", "Hispanic"))),
  LOGCOBS = log(COBS),
  USUBJID = interaction(ID, FORM)
) %>% select(-SEX, -RACE)
```



6. Save the above modifications as a new csv file

```{r}
write_csv(final_data, "iv_oral_alldat.csv", na = ".")
```

## Descriptive Statistics

Objectives

* How to make summaries of the data using descriptive statistics and other data manipulation tools (dplyr, base R functions etc)

### Task III


1. show a summary for all demographic columns

```{r}
final_data <- final_data %>% 
  mutate(GENDER = as.factor(GENDER),
         RACEN = as.factor(RACEN))
uid_final_data <- final_data %>% distinct(ID, .keep_all = TRUE)

uid_final_data %>% 
  select(WT, AGE, GENDER, RACEN) %>%
 summary %>% kable
```


2. Count the number of males/females in the dataset
```{r}
# be careful only 1 row per id if calculating this way
uid_final_data %>% nrow
# or
n_distinct(uid_final_data$ID)
```


3. Count the number of subjects in each "Race" category

```{r}
uid_final_data %>%  
  group_by(RACEN) %>% 
  tally
```

4. calculate the min, mean, and max values for WT, AGE:
    a. by Gender
```{r}
uid_final_data %>% 
  select(GENDER, WT, AGE) %>%
  group_by(GENDER) %>% 
  summarize_all(funs(min, mean, max)) %>% 
  kable
```

    b. by Race
    
```{r}
uid_final_data %>% select(RACEN, WT, AGE) %>%
  group_by(RACEN) %>% 
  summarize_all(funs(min, mean, max)) %>% 
  kable
```

5. What is the Average numbers samples(observations) per individual in this dataset. Hint: make sure you are *only* counting samples, not necessarily all rows are observations!

```{r}
# don't want dosing observations
final_data %>% filter(is.na(AMT)) %>% group_by(ID) %>% 
  summarize(num_obs = n()) %>%
  summarize(avg_samples = mean(num_obs))
```


6. Calculate the Mean, 5th, and 95th percentile concentration at each time point for each formulation and dose level. hint: you can use `?quantile` to calculate various quantiles

```{r}
final_data %>%
  group_by(TIME) %>% 
  s_quantiles(COBS, probs = c(0.05, 0.5, 0.95)) %>% 
  kable
```


