---
title: "data manipulation hands-on"
author: "Devin"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

```{r}
library(knitr)
library(tidyverse)
library(PKPDmisc)
```

## DATA IMPORT


Objectives:

* Import datasets and documents
* Perform basic data manipulation upon importing the data.

```{r}
pk_data <- read_csv("../data/pk_data.csv")
```

```{r}
head(pk_data)
```


## DATA MANIPULATION

The goals of this section:

* Use data manipulation tools to prepare the dataset for analysis


1.  Rename "DV" column as "COBS"

```{r}
pk_data_cobs <- pk_data %>% rename(COBS = DV)
```


4. Perform the following tasks:

a. Ensure that the following columns are numeric and not text: TIME, COBS, WT, AGE, AMT and DOSEs
    
```{r}
glimpse(pk_data_cobs)
```

```{r}
unique_non_numerics(pk_data_cobs$COBS)
```


b. Create a new column called BQLFLAG which takes a value of 
`0` if there is a numerical value in CObs and `1` if there is "BQL" in COBS.
    
```{r}
pk_data_cobs <- pk_data_cobs %>% 
  mutate(BQLFLAG = ifelse(COBS == "BQL", 1, 0),
        NONNUMERICS = ifelse(COBS %in% unique_non_numerics(COBS), 1, 0),
        COBS = as_numeric(COBS))
```

    c. Create a new column called "GENDER" where:
        i. Female = 0
        ii. Male = 1 
    d. Create a new column called RACEN where:
        i. Caucasian = 0
        ii. Asian = 1
        iii. Black = 2
        iv. Hispanic = 3

```{r}
pk_data_cobs <- pk_data_cobs %>%
  mutate(
      GENDER = factor(SEX, 
                      levels = c(0, 1), 
                      labels = c("Female", "Male")),
      RACEN = ifelse(RACE == "Caucasian", 0,
                       ifelse(RACE == "Asian", 1,
                              ifelse(RACE == "Black",2,
                                     ifelse(RACE == "Hispanic", 3, -99))))) 
```

f. Create a new column called "IDF" - unique subject ID as combination of formulation and ID 

```{r}
pk_data_cobs <- pk_data_cobs %>% mutate(IDF = paste(ID, FORM, sep = "-"))
```

g. Remove the following columns
    i. SEX
    ii. RACE
   
```{r}
pk_data_output <- pk_data_cobs %>% select(-SEX, -RACE)
```

```{r}
head(pk_data_output) %>% kable()
```

Save the above modifications as a new csv file

```{r}
write_csv(pk_data_output, "../data/pk_data_output.csv")
```

## Descriptive Statistics

1. show a summary for all demographic columns

```{r}
# single row per id
sid_pk_data <- pk_data_cobs %>%
  distinct(ID, .keep_all = TRUE) 

sid_pk_data %>%
  select(WT, AGE, RACE, SEX) %>%
  mutate(RACE = as.factor(RACE),
         SEX = as.factor(SEX)) %>%
  summary
```

2. Count the number of subjects in each "Race" category

```{r}
# fastest
sid_pk_data %>% count(RACE) %>% kable()
```

```{r eval = F}
# more manual
sid_pk_data %>%
  group_by(RACE) %>%
  tally

# most manual but have more control over column names
sid_pk_data %>%
  group_by(RACE) %>%
  summarize(num_per_race = n()) %>%
  kable()
```


3. calculate the min, mean, and max values for WT, AGE:
    a. by Sex

```{r}
sid_pk_data %>%
  group_by(SEX) %>%
  summarize(
    WT_min = min(WT),
    WT_mean = mean(WT),
    WT_max = max(WT),
    AGE_min = min(AGE),
    AGE_mean = mean(AGE),
    AGE_max = max(AGE)
  ) %>%
  kable()
```

there are also targeted verbs in the form `<verb>_at` that can specify what
columns to act on, and which functions to run

```{r}
sid_pk_data %>%
  group_by(SEX) %>%
  summarize_at(vars(WT, AGE), funs(min, mean, max)) %>%
  kable()
```

    
4. What is the Average numbers samples(observations) per individual in this dataset. 
Hint: make sure you are *only* counting samples, rows with AMT values are not considered observations.

```{r}
# observations are those with NA AMT values
pk_data_cobs %>%
  filter(is.na(AMT)) %>%
  group_by(ID) %>%
  summarize(num_obs = n()) %>%
  # ungroup so no longer calculating by grouping variable ID
  ungroup %>%
  summarize(average_obs = mean(num_obs)) %>% kable()
```


5. Calculate the Mean, 5th, and 95th percentile concentration at each time point for each formulation and dose level. hint: you can use `?quantile` to calculate various quantiles

```{r}
pk_data_cobs %>%
    mutate(COBS = as.numeric(COBS)) %>%
    filter(!is.na(COBS)) %>%
    group_by(TIME, FORM, DOSE) %>%
    summarize(q05 = quantile(COBS, 0.05),
              q50 = quantile(COBS, 0.5),
              q95 = quantile(COBS, 0.95)) %>%
  arrange(FORM, DOSE, TIME) %>% 
  head %>% kable()
```

```{r}
devtools::session_info()
```

