# Diagnostic Plots

1) read in the csv datasets:

* EtaCov_gathered
* Residuals
* Theta

```{r}
library(PKPDmisc)
library(knitr)
library(tidyverse)
```

```{r}
resid <- read_phx("../data/Residuals.csv")
theta <- read_phx("../data/Theta.csv")
etacov_gathered <- read_phx("../data/EtaCov_gathered.csv")
```

2) From the Theta table, create a reasonable quality output table of the results.
Hint, use knitr::kable, in combination with results='asis' in the chunk settings

requires names:

```{r}
theta %>% 
  select(-one_of(c("Scenario", "Var. Inf. factor"))) %>% 
  kable(digits = 2)
```


* clean up columns
* clean up column names
* units

3) Create a CWRES vs Time plot with loess fits for the central tendency and the spread (hint abs() is your friend for the spread)

```{r}
gg_cwres_tad <- function(df) {
df %>%
  ggplot(aes(x = TAD, y = CWRES)) + geom_point() +
  stat_smooth(method = "loess", se=F, color = "red") +
  stat_smooth(data = df %>%
                mutate(CWRES = abs(CWRES)), 
              se = F, color = "blue") +
  stat_smooth(data = df %>%
                mutate(CWRES = -abs(CWRES)), 
              se = F, color = "blue") +
    theme_bw() +
    base_theme()
}

```



```{r}
gg_cwres_tad(resid)
```





4) update the CWRES vs Time plot to flag anything with CWRES > 2.5 as a red value
```{r}

resid %>% 
  mutate(HIGHCWRES = ifelse(abs(CWRES) > 2.5, 1, 0)) %>%
    ggplot(aes(x = TAD, y = CWRES)) +
  geom_point(aes(color = factor(HIGHCWRES))) +
  scale_color_manual(values = c("black", "red"), name = "Outlier", labels = c("not outlier", "outlier")) +
  stat_smooth(method = "loess") +
  stat_smooth(data = resid %>%
                mutate(CWRES = abs(CWRES)), 
              method="loess", color = "red", se = F) +
  stat_smooth(data = resid %>%
                mutate(CWRES = -abs(CWRES)), 
              method="loess", color = "red", se = F) 

```

5) print a table of key information for all points with CWRES > 2.5

```{r}
resid %>% 
  mutate(HIGHCWRES = ifelse(abs(CWRES) > 2.5, 1, 0)) %>%
  filter(HIGHCWRES ==1) %>% select(ID, IVAR, TAD, IPRED, DV) %>% kable(digits = 2)
```


6) Plot individual IPRED and DV vs time

```{r}
split_resid <- resid %>% filter(TADSeq ==1) %>% mutate(IDBINS = ids_per_plot(ID, 9)) %>% split(.[["IDBINS"]])

p <- function(df) {
  df %>%
  ggplot(aes(x = TAD, y = IPRED, group= TADSeq)) +
  geom_line() + facet_wrap(~ID) + theme_bw() +
    geom_point(aes(x = TAD, y = DV))+
    labs(list(x = "Time after Dose, hrs",
              y = "Individual Predicted and Observed")) 
}
split_resid %>% map(p)
```

As a reminder, map works like lapply, it applies the same function to each element in the list. In this case, it is taking split_resid (which is the residual dataframe split by 9 ids per group) and then applies the plot function to each set of 9.


6b) add the population prediction as a dashed blue line

```{r}

p <- function(df) {
  df %>%
  ggplot(aes(x = TAD, y = IPRED, group= TADSeq)) +
  geom_line() + facet_wrap(~ID) + theme_bw() +
    geom_point(aes(x = TAD, y = DV))+labs(list(x = "Time after Dose, hrs", y = "Individual Predicted and Observed")) +
    geom_line(aes(x = TAD, y = PRED, group = TADSeq), color = "blue")
}

split_resid %>% map(p)
```

7) With EtaCov_final create histograms of all the eta distributions

```{r}

p_etas<- etacov_gathered %>%
  ggplot(aes(x = VALUE, group = ETA)) + 
  geom_histogram(fill = "white", color = "black") + 
  facet_wrap(~ETA, scales = "free") + base_theme()

p_etas
```

add a mean value for each eta overlaid on the above plot
```{r}
mean_eta <- etacov_gathered %>% 
    group_by(ETA) %>%
  summarize(meanEta = mean(VALUE))

p_etas + 
  geom_vline(data = mean_eta, aes(xintercept = meanEta), size = 1.5, color = "red")
```

8) Create Eta vs Covariate plots for each covariate and all etas

```{r}
etacov_gathered %>%
    ggplot(aes(x = WT, y = VALUE, group = ETA)) + 
  geom_point() + facet_wrap(~ETA, scales = "free") + 
  stat_smooth(method = "loess", color = "blue", se = F, size = 1.3) + 
  base_theme()

etacov_gathered %>%
    ggplot(aes(x = AGE, y = VALUE, group = ETA)) + 
  geom_point() + facet_wrap(~ETA, scales = "free") +
  stat_smooth(method = "loess", color = "blue", se = F, size = 1.3) + base_theme()
```

Note in the plot above, the choice of facet_wrap was arbitrary, and potentially a cleaner looking plot can be created with facet_grid, especially for labels, my suggestion is to try both.

Hint: since there is so much duplicated, this would be a good opportunity to turn that into a function that you pass in the covariate to plot for `x`.

9) add loess fits to the eta cov plots

done in above plots
