# Diagnostic Plots

1) read in the csv datasets:

* Residuals

```{r}
library(PKPDmisc)
library(knitr)
library(tidyverse)
```

```{r}
resid <- read_csv("../data/Residuals.csv")
```

* Create a Res vs Time function with loess fits for the central tendency and the spread (hint abs() is your friend for the spread). 
    * Conditionally allow the loess curve of central tendency to appear, 
    with a default of TRUE. 
    * Users should be able to specify the residual column name.

```{r}
gg_res_tad <- function(df, .tad, .res, .show_loess = TRUE) {
    .tad <- rlang::enexpr(.tad)
    .res <- rlang::enexpr(.res)
    ple <- rlang::quo(
        df %>%
          ggplot(aes(x = !!.tad, y = !!.res)) + geom_point() +
          stat_smooth(data = df %>%
                        mutate(!!.res := abs(!!.res)), 
                      se = F, color = "blue") +
          stat_smooth(data = df %>%
                        mutate(!!.res := -abs(!!.res)), 
                      se = F, color = "blue") +
            theme_bw()
    )
    output <- rlang::eval_tidy(ple) 
    if (.show_loess) {
       return(
           output + 
            stat_smooth(method = "loess", se=F, color = "red")
       ) 
    }
    return(output)
}
```


### CWRES vs time after dose (TAD)

```{r}
gg_res_tad(resid, TAD, CWRES)
```

* update your function to flag any point over some threshold as red, with a default of absolute difference of > 2.5 

```{r}
gg_res_tad <- function(df, .tad, .res, .threshold = 2.5, .show_loess = TRUE) {
    .tad <- rlang::enexpr(.tad)
    .res <- rlang::enexpr(.res)
    ple <- rlang::quo(
     df %>% 
        mutate(HIGHRES__ = ifelse(abs(!!.res) > .threshold, 1, 0)) %>%
          ggplot(aes(x = !!.tad, y = !!.res)) +
          geom_point(aes(color = factor(HIGHRES__))) +
          scale_color_manual(values = c("black", "red"), 
                             name = "Outlier", 
                             labels = c("not outlier", "outlier")) +
          stat_smooth(data = df %>%
                        mutate(!!.res := abs(!!.res)), 
                      se = F, color = "blue") +
          stat_smooth(data = df %>%
                        mutate(!!.res := -abs(!!.res)), 
                      se = F, color = "blue") +
            theme_bw()
    )
    output <- rlang::eval_tidy(ple) 
    if (.show_loess) {
       return(
           output + 
            stat_smooth(method = "loess", se=F, color = "red")
       ) 
    }
    return(output)
}
```

```{r}
gg_res_tad(resid, TAD, CWRES)
```

```{r}
devtools::session_info()
```

