## functional techniques



```{r}
resid <- read_csv("../data/Residuals.csv")
eta_cov <- read_csv("../data/EtaCov_base.csv")
```

```{r}
split_resid <- resid %>% filter(TADSeq ==1) %>% mutate(IDBINS = ids_per_plot(ID, 9)) %>% split(.[["IDBINS"]])

p <- function(df) {
  df %>%
  ggplot(aes(x = TAD, y = IPRED, group= TADSeq)) +
  geom_line() + facet_wrap(~ID) + theme_bw() +
    geom_point(aes(x = TAD, y = DV))+
    labs(list(x = "Time after Dose, hrs",
              y = "Individual Predicted and Observed")) 
}
split_resid %>% map(p)
```

As a reminder, map works like lapply, it applies the same function to each element in the list. In this case, it is taking split_resid (which is the residual dataframe split by 9 ids per group) and then applies the plot function to each set of 9.


6b) add the population prediction as a dashed blue line

```{r}

p <- function(df) {
  df %>%
  ggplot(aes(x = TAD, y = IPRED, group= TADSeq)) +
  geom_line() + facet_wrap(~ID) + theme_bw() +
    geom_point(aes(x = TAD, y = DV))+labs(list(x = "Time after Dose, hrs", y = "Individual Predicted and Observed")) +
    geom_line(aes(x = TAD, y = PRED, group = TADSeq), color = "blue")
}

split_resid %>% map(p)
```

7) With EtaCov_base create histograms of all the eta distributions

```{r}

p_etas<- etacov_gathered %>%
  ggplot(aes(x = VALUE, group = ETA)) + 
  geom_histogram(fill = "white", color = "black") + 
  facet_wrap(~ETA, scales = "free") + base_theme()

p_etas
```

add a mean value for each eta overlaid on the above plot
```{r}
mean_eta <- etacov_gathered %>% 
    group_by(ETA) %>%
  summarize(meanEta = mean(VALUE))

p_etas + 
  geom_vline(data = mean_eta, aes(xintercept = meanEta), size = 1.5, color = "red")
```

8) Create Eta vs Covariate plots for each covariate and all etas

```{r}

```

Note in the plot above, the choice of facet_wrap was arbitrary, and potentially a cleaner looking plot can be created with facet_grid, especially for labels, my suggestion is to try both.

Hint: since there is so much duplicated, this would be a good opportunity to turn that into a function that you pass in the covariate to plot for `x`.

9) add loess fits to the eta cov plots

done in above plots

```

