# Introduction to ggplot2

```{r}
library(PKPDmisc)
library(PKPDdatasets)
library(tidyverse)
```

## XY Plots
1) For concentration-time plots filter only OCC 1 from dapa_IV_oral (in PKPDdatasets) for subsequent plotting
```{r}
data1 <- dapa_IV_oral %>% filter(OCC == 1)
```

2) Basic Concentration-Time plot (point and lines)

```{r}
data1 %>% ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point() + geom_line()
```

3) make points/lines bigger

```{r}
data1 %>% ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1)
```

4) add log transformation to y axis

```{r}
data1 %>% ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10()
```

5) Add color by gender

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID, color = GENDER)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10()
```

BONUS: rename labels legend

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID, color = GENDER)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10() + scale_color_discrete(labels = c("Male", "Female"))
```

BONUS: move legend to top right

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID, color = GENDER)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10() + scale_color_discrete(labels = c("Male", "Female")) +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1))
```

6) facet by race

Note, with facetting by race, the color becomes unnessary extra visual stimulus so we remove the color.

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
    facet_wrap(~GENDER)
```

BONUS: rename facet strips by name

There are a couple ways of doing this:

1) on the fly mutate the the values to labels

```{r}
# this is the only time you will see assigning to a column directly
# in this workshop, we encourage to always use mutate to manage columns
# in a dataframe, however as it has not been formally introduced we're
# using base R here
data1$GENDER <-  factor(data1$GENDER, 
                         levels = c(0, 1), 
                         labels = c("Male", "Female")) 
data1 %>%
    ggplot(aes(x = TIME, y = COBS, group = ID)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
    facet_wrap(~GENDER)
```

Use facet_grid with the labeller argument

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
  facet_grid(.~GENDER, labeller = label_both)
```

9) color by weight

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID, color = WEIGHT)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1))
```

BONUS: rename axes

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID, color = WEIGHT)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), 
          legend.justification = c(1, 1)) +
  labs(x = "Time, hours", y = "Concentration, ug/mL")
```


## Boxplots and Histograms

* Histogram(s) of demographics

```{r}
# distinct is a dplyr verb - easiest way to subset data by the 
# first row in some combination.
sid_data <- data1 %>% 
    distinct(ID, .keep_all = TRUE) # single row per id data

sid_data %>%
ggplot(aes(x = WEIGHT)) + 
  geom_histogram(binwidth= 4, color="black", fill="white")
```

* add vertical line for median value
  
```{r}
ggplot(data = sid_data, 
       aes(x = WEIGHT)) + 
  geom_histogram(binwidth= 4, color="black", fill="white") +
  geom_vline(aes(xintercept = median(WT)), size= 2, color = "red")
```


```{r}
devtools::session_info()
```

