# Introduction to ggplot2

```{r}
library(PKPDmisc)
library(PKPDdatasets)
library(tidyverse)
```

## XY Plots
1) For concentration-time plots filter only OCC 1 from dapa_IV_oral (in PKPDdatasets) for subsequent plotting
```{r}
data1 <- dapa_IV_oral %>% filter(OCC == 1)
```

2) Basic Concentration-Time plot (point and lines)

```{r}
data1 %>% ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point() + geom_line()
```

3) make points/lines bigger

```{r}
data1 %>% ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1)
```

4) add log transformation to y axis

```{r}
data1 %>% ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10()
```

5) Add color by gender

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID, color = GENDER)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10()
```

BONUS: rename labels legend

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID, color = GENDER)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10() + scale_color_discrete(labels = c("Male", "Female"))
```

BONUS: move legend to top right

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID, color = GENDER)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10() + scale_color_discrete(labels = c("Male", "Female")) +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1))
```

6) facet by race

Note, with facetting by race, the color becomes unnessary extra visual stimulus so we remove the color.

```{r}
data1 %>% 
    ggplot(aes(x = TIME, y = COBS, group = ID)) +
    geom_point(size = 2.5) + geom_line(size = 1.1) +
    scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
    facet_wrap(~GENDER)
```

BONUS: rename facet strips by name

There are a couple ways of doing this:

1) on the fly mutate the the values to labels

```{r}
data1 %>% 
  mutate(GENDER = factor(GENDER, 
                         levels = c(0, 1), 
                         labels = c("Male", "Female"))) %>%
           ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
  facet_wrap(~GENDER)
```

Use facet_grid with the labeller argument

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
  facet_grid(.~GENDER, labeller = label_both)
```

*note: the newest version of ggplot2 has changed the labeller behavior, when you have a factor that you want to
display both the underlying value and the label, you should now use `labeller = labe_both`, as shown below
7) add mean conc-time profile and color red
```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
  facet_grid(.~GENDER, labeller = label_both) +
  geom_line(data = data1 %>% group_by(TIME) %>%
              summarize(meanCONC = mean(COBS)), 
            aes(x = TIME, y = meanCONC, group = NULL), size = 1.25, color = "red")
```

8) add two mean profiles (one for each gender)

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
  facet_grid(.~GENDER, labeller = label_both) +
  geom_line(data = data1 %>% 
              group_by(TIME, GENDER) %>%
              summarize(meanCONC = mean(COBS)), 
            aes(x = TIME, 
                y = meanCONC, 
                group = GENDER, 
                color = GENDER), 
            size = 1.25) + 
    scale_color_discrete(name = "Mean Profile", labels = c("Male", "Female"))
```

9) color by weight

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID, color = WEIGHT)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1))
```

BONUS: rename axes

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID, color = WEIGHT)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
  xlab("Time, hours") + ylab("Concentration, ug/mL")
# alternative to xlab/ylab is declare all at once
# labs(list(x = "Time, hours", y = "Concentration, ug/mL"))
```

BONUS: additional customizations

```{r}
data1 %>% 
   ggplot(aes(x = TIME, y = COBS, group = ID, color = WEIGHT)) +
  geom_point(size = 2.5) + geom_line(size = 1.1) +
  scale_y_log10() +
    theme(legend.position = c(1, 1), legend.justification = c(1, 1)) +
  xlab("Time, hours") + ylab("Concentration, ug/mL") +
  # all these theme customizations are defined in PKPDmisc::base_theme()
  # however I wanted to show what the individual customizations looked like
  # so you can play around and test various
theme(legend.text = element_text(size = 10),
     legend.title = element_text(size = 12),
     axis.title.x = element_text(size = 12, face = "bold"),
     axis.title.y = element_text(size = 12, face = "bold"),
     axis.text.x = element_text(color = "black", size = 10),
     axis.text.y = element_text(color = "black", size = 10))
```

## Boxplots and Histograms
* Histogram(s) of demographics
```{r}
sid_data <- data1 %>% distinct(ID, .keep_all = TRUE) # single row per id data

sid_data %>%
ggplot(aes(x = WEIGHT)) + 
  geom_histogram(binwidth= 4, color="black", fill="white")
```

  * add vertical line for median value
```{r}
 median <- sid_data %>% summarize(medianWT = median(WEIGHT))
 # alternative explicit declaration of data inside ggplot rather than outside using %>%
ggplot(data = sid_data, 
       aes(x = WEIGHT)) + 
  geom_histogram(binwidth= 4, color="black", fill="white") +
  geom_vline(xintercept = median[["medianWT"]], size= 2, color = "red")
```

* Boxplot cmax/AUC by gender
```{r}
cmaxauc <- data1 %>% group_by(ID, GENDER) %>% 
    summarize(cmax = max(COBS), aucinf = auc_inf(TIME, COBS))
ggplot(cmaxauc, aes(x = GENDER, y = cmax, group = GENDER)) + geom_boxplot()
ggplot(cmaxauc, aes(x = GENDER, y = aucinf, group = GENDER)) + geom_boxplot()
```

```{r}
devtools::session_info()
```

