# Tidying Data

```{r}
library(knitr)
library(tidyverse)
library(PKPDmisc)
```

```{r}
ebe_cov <- read_csv("../data/ebe_cov.csv")
```

Remove all the columns with ETAS and name dataframe `covs`

```{r}
covs <- ebe_cov %>% select(-contains("ETA"))
```

Find all columns with an NA value, and the associated ID

```{r}
has_missing <- covs %>% 
  group_by(ID) %>%
  select_if(~any(is.na(.))) %>%
  ungroup()
head(has_missing)

missing_ids <- has_missing %>%
  gather(cov, values, -ID) %>%
  filter(is.na(values))
missing_ids
```

Impute the NA values using the mean for any continous variable, and largest group for categorical

```{r }
has_missing %>%
  select(-ID) %>%
  mutate(SEX = as.factor(SEX)) %>% summary
```

```{r}
replacement_values <- has_missing %>% 
  summarize(BW = mean(BW, na.rm = T)) %>% 
  mutate(SEX = 0)
```

```{r}
covs <- covs %>% 
  replace_na(replace = as.list(replacement_values)) 
```


```{r }
covs %>%
  filter(ID %in% missing_ids$ID)
```

## data checkout for all covariates

Set all categorical covariates to factors with appropriate labels

```{r}
covs_f <- covs %>% 
  mutate(SEXC = factor(SEX, 
                       levels = c(0, 1), 
                       labels = c("FEMALE", "MALE")
                       ),
        RACEC = factor(RACE, 
                       levels = c(0:2), 
                       labels = c("WHITE", "BLACK", "ASIAN")
                       )
         ) %>%
  select(-SEX, -RACE)
```

```{r}
g_cont_covs <- covs_f %>% gather(cov, value, BW:CRCL)
```

Plot a scatter plot of all continuous covariates versus ID to check for visual outliers

```{r}
g_cont_covs %>%
  ggplot(aes(x = ID, y = value)) +
  geom_point() +
  facet_wrap(~cov, scales = "free")
```

Plot a violin/box plot of all continuous covariates versus SEX to check for visual trends 

```{r}
g_cont_covs %>%
  ggplot(aes(x = SEXC, y = value)) +
  geom_violin() +
  geom_jitter(width = 0.1) +
  facet_wrap(~cov, scales = "free")
```

Plot a violin/box plot of all continuous covariates versus all categorical covariates to check for visual trends 

```{r fig.height = 10, fig.width = 8}
g_cont_covs %>%
  gather(catcov, catvals, RACEC, SEXC) %>%
  ggplot(aes(x = catvals, y = value)) +
  geom_violin() +
  geom_jitter(width = 0.1) +
  facet_grid(cov~catcov, scales = "free")
```

```{r}
devtools::session_info()
```

